Pattern matching for\\ $(213,231)$-avoiding permutations

### 2015-02-09 article.tex
add : lemma and corollary for section 4 to explain a safe match

%Pattern Marching

\DeclareMathOperator{\PMa}{PM}
\newcommand{\PM}[6]{\PMa_{{#1}}^{{#2},{#3},{#4}}({#5},{#6})}

@@

@@

\begin{lemma}
\label{lemma:MatchStripeToPermutation}
Let $\pi$ and $\sigma$ be two $(213,231)$-avoiding permutations,
and $t$ be a subsequence of $\pi$ of length $\sigma$.
Then $t$ is an occurrence of $\sigma$ in $\pi$ if and only if
$\stripew{t}$ is an occurrence of $\stripew{\sigma}$ in $\stripew{\pi}$.
\end{lemma}

\begin{proof}
	The forward direction is obvious.
	We prove the backward direction by induction on the size of the motif
	$\sigma$.
	The base case is a motif of size 2. Suppose that we have the motif $(12)$, thus the stripe of the motif is $\ustep$. Let's $t$ be a subsequence of $\ptext$ such that $\stripew{t}$ is a occurrence of $\stripew{\pmotif}$ in $\stripew{\ptext}$. Because we have $\stripe{t}{1}=\ustep$, the next element of $t[1]$ is $\ptext$ is superior to  $t[1]$, so we have a match. Same idea goes of motif $(21)$.
	Now, suppose that the lemma is true for every motif in $\Av{n}{(231,213)}$. Suppose that we have a subsequence $t$ of $\ptext$ such that $\stripew{t}$ is an occurrence of $\stripew{\pmotif}$ in $\stripew{\ptext}$, then $\stripew{t[2:]}$ is an occurrence of $\pmotif[2:]$ in $\ptext$. Moreover
	$\pmotif[2:] \in \Av{n}{(231,213)}$ so by hypothesis of induction $t[2:]$ is an occurrence of $\sigma[2:]$ in $\ptext$.
	Moreover if $\stripe{\pmotif}{1}$ is ascend (descent), then by lemma \ref{lemma:first element is 1 or n} $\pmotif[1]$ is the minimal (maximal) element of the motif. Furthermore $\stripew{t}$ is an occurrence of $\stripew{\pmotif}$ in $\stripew{\ptext}$ , so $\stripe{t}{1}$ is also ascend (descent) thus $t[1]$ is also the minimal (maximal) element of the subsequence $t$ which complete our proof.

\qed
\end{proof}

@@

@@

\begin{proof}%[of Proposition~\ref{Proposition:sigma avoids 213 and 231}]
Given an ascent (descent) factor $\factor{i}$ of a motif $\pmotif$, if we want to match the suffix of $\pmotif$ starting by $\factor{i}$, $\factor{i}\factor{i-1}\cdots\factor{1}$ and given that we know every match of $\factor{i-1}\cdots\factor{1}$, the optimal match is to choose the match of $\factor{i-1}\cdots\factor{1}$ that minimize (maximize) the maximal (minimal) element of the match of $\factor{i-1}\cdots\factor{1}$ and that allow a match of factor $\factor{i}$. In other word if we want to match $\pmotif[\first{i}:]$ with $\ptext[j:]$ such that $\first{i}$ is matched with element at index $j$ in $\ptext$, the optimal way of matching $\pmotif[\first{i-1}:]$ is to choose the minimal (maximal) element at index $j'$ of $\ptext[j:]$ such that : (1) $\pmotif[\first{i-1}:]$ is matched with $\ptext[j':]$ and element $\first{i-1}$ is matched with element at index $j'$, (2) the factor $\factor{i}$ is matched with $\ptext[j,j'-1]$ and element $\first{i}$ is matched with element at index $j$, (3) Every elements of the match of (2) if inferior (superior) to the minimal (maximal) element of the match of (1). Indeed when we will match $\factor{i+1}\factor{i}\factor{i-1}\cdots\factor{1}$, every element of $\factor{i+1}$ must be matched to element superior (inferior) to every element matched element of $\factor{i}\factor{i-1}\cdots\factor{1}$, particularly its must be superior (inferior) to the maximal (minimal) element of the match of $\factor{i}\factor{i-1}\cdots\factor{1}$.

\begin{remark}
Given an ascent (descent) factor $\factor{i}$, the maximal (minimal) element of $\factor{i}\factor{i-1}\cdots\factor{1}$ is $\first{i-1}$.
\end{remark}

\begin{remark}
An ascent (descent) factor 	$\factor{i}$ is a match in $\ptext[j:j']$  if and only if there exists an increasing (decreasing) subsequence in $\ptext[j:j']$ of size superior or equal to the size of $\factor{i}$.
\end{remark}

Consider the following problem :\\
$LM(i,j)$ = the optimal match of $\first{i-1}$
	if there exists a match of $\pmotif[\first{i}:]$
	with $\ptext[j:]$
	and element $\first{i}$
	is matched with element at index $j$.\\

@@

With LIS(i,j,k) (LDS(i,j,k)) is the longest increasing (decreasing) sequence in $\ptext$ starting at $i$ and ending at $j$,
with every element of this sequence
inferior (superior) to $k$.
LIS and LDS can be computed in $O(|\ptext|^2*log(log(|\ptext|)))$ (see \cite{Bespamyatnikh00enumeratinglongest}).\\

@@

	\begin{proposition}
	Given a permutation $\pi$,
	finding the longest subsequence avoiding (231,213)
	can be done in $O(|\ptext|log(log(|\ptext|)))$ time and in $O(n)$ space.
	\end{proposition}

	\begin{proof}
	The proposition \ref{proposition:longestIncreasingSubsequence} lead to algorithm
	where we have to compute longest increasing and decreasing subsequence ending at every index. Then finding the maximum sum of longest increasing and decreasing subsequence ending at the same index.
	Computing the longest increasing and decreasing can be done in $O(|\ptext|*log(log(|\ptext|)))$ time and $O(n)$ space (see \cite{Bespamyatnikh00enumeratinglongest}), then finding the maximum can be done in linear time.
	\end{proof}

>>>>

%Pattern Marching
\DeclareMathOperator{\PMa}{PM}
\newcommand{\PM}[6]{\PMa_{{#1}}^{{#2},{#3},{#4}}({#5},{#6})}

@@

%SET OF OCCURENCE
\DeclareMathOperator{\SETa}{S}
\newcommand{\SET}[4]{\SETa_{{#1}}^{{#2}}({#3},{#4})}

@@

@@

To ease the lemma and its proof we define the set $\SET{\pmotif}{\ptext}{i}{j}$ such that if $s \in S$, then $s$ is an occurrence of $\factor{i} \factor{i-1} \cdots \factor{1}$ in $\ptext[j:]$ and $s[\first{i}]=\ptext[j]$.

\begin{lemma}
Given a permutation $\pmotif$ and $\factor{i}$ an ascent (descent) factor.
Let's choose a subsequence $s \in \SET{\pmotif}{\ptext}{i}{j}$, such that $s[\first{i-1}]$ is minimal (maximal), given a subsequence
$s' \in \SET{\pmotif}{\ptext}{i}{j}$ and a subsequence $t=t's'$ of $\ptext$  if $t$ is an occurrence of $\factor{i+1} \factor{i} \factor{i-1} \cdots \factor{1}$ then the subsequence $t's$ is an occurrence of $\factor{i+1} \factor{i} \factor{i-1} \cdots \factor{1}$ in $\ptext$.
\end{lemma}

\begin{proof}
Given a permutation $\pmotif$ and $\factor{i}$ an ascent (descent) factor.
Let's choose a subsequence $s \in \SET{\pmotif}{\ptext}{i}{j}$, such that $s[\first{i-1}]$ is minimal (maximal).
Let's $s' \in \SET{\pmotif}{\ptext}{i}{j}$ and a subsequence $t=t's'$ of $\ptext$ such that $t$ is an occurrence of $\factor{i+1}\factor{i}\factor{i-1}\cdots\factor{1}$. 
Remark that every element of $t'$ must be superior (inferior) to every element of $s'$, and that the maximal element of $s'$ is the element matched to $t[\first{i-1}]$. Thus if we have a subsequence $s \in \SET{\pmotif}{\ptext}{i}{j}$ such that $s[\first{i-1}]$ is minimal (maximal), then $s[\first{i-1}]\leq s'[\first{i-1}]$, so every element of $'t$ is also superior (inferior) to every element of $s$. Thus the subsequence $t's$ is an occurrence of $\factor{i+1}\factor{i}\factor{i-1}\cdots\factor{1}$ in $\ptext$.
\end{proof}

\begin{corollary}
Let's choose a subsequence $s \in \SET{\pmotif}{\ptext}{i}{j}$, such that $s[\first{i-1}]$ is minimal (maximal). There is a match of $\pmotif$ in $\ptext$ with element $\first{i}$ matched to $\ptext[j]$ if and only if there exists $t$ such that $ts$ is a match of $\pmotif$ in $\ptext$. 
\end{corollary}

This corollary give us a safe way to choose a match.

@@

\begin{proof}%[of Proposition~\ref{Proposition:sigma avoids 213 and 231}]
%Given an ascent (descent) factor $\factor{i}$ of a motif $\pmotif$, if we want to match the suffix of $\pmotif$ starting by $\factor{i}$, $\factor{i}\factor{i-1}\cdots\factor{1}$ and given that we know every match of $\factor{i-1}\cdots\factor{1}$, the safest match is to choose the match of $\factor{i-1}\cdots\factor{1}$ that minimize (maximize) the maximal (minimal) element of the match of $\factor{i-1}\cdots\factor{1}$ and that allow a match of factor $\factor{i}$. Indeed when we will match $\factor{i+1}\factor{i}\factor{i-1}\cdots\factor{1}$, every element of $\factor{i+1}$ must be matched to element superior (inferior) to every element matched element of $\factor{i}\factor{i-1}\cdots\factor{1}$, particularly its must be superior (inferior) to the maximal (minimal) element of the match of $\factor{i}\factor{i-1}\cdots\factor{1}$. In other word if we want to match $\pmotif[\first{i}:]$ with $\ptext[j:]$ such that $\first{i}$ is matched with element at index $j$ in $\ptext$, the optimal way of matching $\pmotif[\first{i-1}:]$ is to choose the minimal (maximal) element at index $j'$ of $\ptext[j:]$ such that : (1) $\pmotif[\first{i-1}:]$ is matched with $\ptext[j':]$ and element $\first{i-1}$ is matched with element at index $j'$, (2) the factor $\factor{i}$ is matched with $\ptext[j,j'-1]$ and element $\first{i}$ is matched with element at index $j$, (3) Every elements of the match of (2) if inferior (superior) to the minimal (maximal) element of the match of (1).
%
%\begin{remark}
%Given an ascent (descent) factor $\factor{i}$, the maximal (minimal) element of $\factor{i}\factor{i-1}\cdots\factor{1}$ is $\first{i-1}$.
%\end{remark}
%
%\begin{remark}
%An ascent (descent) factor 	$\factor{i}$ is a match in $\ptext[j:j']$  if and only if there exists an increasing (decreasing) subsequence in $\ptext[j:j']$ of size superior or equal to the size of $\factor{i}$.
%\end{remark}

Consider the following problem :\\
$LM(i,j)$ = the safest match of $\first{i-1}$
	if there exists a match of $\pmotif[\first{i}:]$
	with $\ptext[j:]$
	and element $\first{i}$
	is matched with element at index $j$.\\

@@

With LIS(i,j,k) (LDS(i,j,k)) is the longest increasing (decreasing) sequence in $\ptext$ starting at $i$ and ending at $j$,
with every element of this sequence
inferior (superior) to $k$.
LIS and LDS can be computed in $O(|\ptext|^2*\log(\log(|\ptext|)))$ (see \cite{Bespamyatnikh00enumeratinglongest}).\\

@@

	\begin{proposition}
	Given a permutation $\pi$,
	finding the longest subsequence avoiding (231,213)
	can be done in $O(|\ptext|\log(\log(|\ptext|)))$ time and in $O(n)$ space.
	\end{proposition}

	\begin{proof}
	The proposition \ref{proposition:longestIncreasingSubsequence} lead to algorithm
	where we have to compute longest increasing and decreasing subsequence ending at every index. Then finding the maximum sum of longest increasing and decreasing subsequence ending at the same index.
	Computing the longest increasing and decreasing can be done in $O(|\ptext|*\log(\log(|\ptext|)))$ time and $O(n)$ space (see \cite{Bespamyatnikh00enumeratinglongest}), then finding the maximum can be done in linear time.
	\end{proof}
